---
title: "Oligogenic Results"
output: html_document
---

## Power Calculation for Causal Definition

In an oligogenic setting, with an infinitesimal background, we cannot simply set our causal variants as those with nonzero effect sizes. Thus, we ran SuSiE with L = 1 on multiple different total PVE settings, each on 150 different LD blocks with n = 1500 and ranging from p = 5000 to p = 8000, and with only a single causal variant. We found at PVE = 0.0225 we can achieve 80% power. 

```{r, echo = F, warning = F, message = F}
library(tidyverse)

# List all .rds files from the directory
files <- list.files(
  "/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/scripts/benchmark/power_calculation",
  pattern = "\\.rds$", full.names = TRUE
)

# Initialize a data frame to store the results
results <- data.frame(h2total = numeric(), CS_Recall = numeric(), stringsAsFactors = FALSE)

# Loop over each file
for (filename in files) {
  # Load the .rds file
  file_data <- readRDS(filename)
  
  # Extract the h2total parameter from the file name.
  # This regex finds the number following "h2total" and before the next underscore.
  h2_value <- as.numeric(sub(".*h2total([0-9.]+)_.*", "\\1", filename))
  
  # Extract the CS_Recall value from the file_data object.
  cs_recall <- file_data$avg_metrics$CS_Recall
  
  # Append the extracted values to the results data frame
  results <- rbind(results, data.frame(h2total = h2_value, CS_Recall = cs_recall))
  
  # Remove the loaded object and run garbage collection to free memory
  rm(file_data)
  gc()
  
  # Inform that the current file has been processed
  message("Processed original file: ", filename)
}

# Create the plot with ggplot2
plot <- ggplot(results, aes(x = h2total, y = CS_Recall)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0.8, linetype = "dotted", color = "red") +
  ylim(0, 1) +
  labs(title = "Power Calculation", x = "Total PVE", y = "CS Power") +
  theme_minimal()

# Display the plot
print(plot)

```

## Recall & FDR Plots

The following plots show the Recall and FDR for our oligogenic simulation setting depending on the heritability level (averaged across 50 replicates). 

- Fineboost was ran using a null_max_threshold of 0.015.
- SuSiE.ash (Grid 2) refers to the SuSiE.ash (Marginal) model with a variance grid estimated from the data.

```{r, echo=FALSE,warning=F,message=F}
# Load required libraries
library(tidyverse)
library(patchwork)
library(ROCR)
library(gridExtra)

# ------------------------------
# Helper Functions
# ------------------------------

# Function to compute credible set metrics (FDR and Recall)
compute_cs_metrics <- function(credible_sets, truth) {
  if (length(credible_sets) == 0) {
    return(list(cs_fdr = 0, cs_recall = 0))
  }
  TP <- sum(sapply(credible_sets, function(cs) any(cs %in% truth)))
  FP <- length(credible_sets) - TP
  cs_fdr <- ifelse((TP + FP) > 0, FP / (TP + FP), 0)
  
  recovered <- unique(unlist(credible_sets))
  TP_recall <- sum(truth %in% recovered)
  FN_recall <- length(truth) - TP_recall
  cs_recall <- ifelse((TP_recall + FN_recall) > 0, TP_recall / (TP_recall + FN_recall), 0)
  
  list(cs_fdr = cs_fdr, cs_recall = cs_recall)
}

# Function to define causal variants using a PVE threshold
is_causal <- function(beta, residual_variance, pve_threshold) {
  variance_explained <- beta^2
  var_g <- sum(variance_explained)
  total_variance <- var_g + residual_variance
  proportion_var_explained <- variance_explained / total_variance
  which(proportion_var_explained > pve_threshold)
}

# Function to compute AUROC (using ROCR)
compute_auroc <- function(truth, pips) {
  pred <- prediction(pips, truth)
  perf <- performance(pred, "auc")
  as.numeric(perf@y.values)
}

# Function to compute AUPRC (using ROCR)
compute_auprc <- function(truth, pips) {
  pred <- prediction(pips, truth)
  perf <- performance(pred, "aucpr")
  as.numeric(perf@y.values)
}

# Set parameters
#pve_threshold <- 0.005
pve_threshold <- 0.0225
top_N <- c(1, 3, 5, 10, 15, 25)

# ------------------------------
# Pre-filter File Lists
# ------------------------------

# For original files, exclude those with h2total = 0.3
orig_files <- list.files(
  "/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/scripts/benchmark/oligo_result_files_original",
  pattern = "\\.rds$", full.names = TRUE
)
orig_files <- orig_files[ !grepl("h2total0\\.3", orig_files) ]

# For new files, only keep those with h2total = 0.15 or 0.50, Klength = 20, and warmStart = 1
new_files <- list.files(
  "/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/scripts/benchmark/oligo_result_files",
  pattern = "warmStart", full.names = TRUE
)
new_files <- new_files[ grepl("h2total(0\\.15|0\\.5)", new_files) &
                           grepl("Klength20", new_files) &
                           grepl("warmStart1", new_files) ]

# ------------------------------
# Process Original Files (CS Metrics)
# ------------------------------

results_list_orig <- list()

for (file in orig_files) {
  filename <- basename(file)
  h2total <- as.numeric(str_extract(filename, "(?<=h2total)[0-9.]+"))
  message("Processing original file: ", filename)
  file_data <- readRDS(file)
  
  for (i in seq_along(file_data$all_results)) {
    rep_data <- file_data$all_results[[i]]
    beta <- rep_data$data$beta
    residual_variance <- rep_data$data$residual_variance
    truth <- is_causal(beta, residual_variance, pve_threshold)
    n_causal <- length(truth)
    
    ## Method 1: SuSiE
    cs_susie <- if (!is.null(rep_data$susie_out$sets$cs)) rep_data$susie_out$sets$cs else list()
    metrics <- compute_cs_metrics(cs_susie, truth)
    results_list_orig[[length(results_list_orig) + 1]] <- data.frame(
      file = filename,
      replicate = i,
      h2total = h2total,
      pve_threshold = pve_threshold,
      n_causal = n_causal,
      Method = "SuSiE",
      cs_fdr = metrics$cs_fdr,
      cs_recall = metrics$cs_recall,
      stringsAsFactors = FALSE
    )
    
    ## Method 2: SuSiE-inf
    cs_susie_inf <- if (!is.null(rep_data$susie_inf_out$sets)) rep_data$susie_inf_out$sets else list()
    metrics <- compute_cs_metrics(cs_susie_inf, truth)
    results_list_orig[[length(results_list_orig) + 1]] <- data.frame(
      file = filename,
      replicate = i,
      h2total = h2total,
      pve_threshold = pve_threshold,
      n_causal = n_causal,
      Method = "SuSiE-inf",
      cs_fdr = metrics$cs_fdr,
      cs_recall = metrics$cs_recall,
      stringsAsFactors = FALSE
    )
    
    ## Method 3: SuSiE-ash (Scenario 6)
    cs_susie_ash <- if (!is.null(rep_data$susie_ash_6$sets)) rep_data$susie_ash_6$sets else list()
    metrics <- compute_cs_metrics(cs_susie_ash, truth)
    results_list_orig[[length(results_list_orig) + 1]] <- data.frame(
      file = filename,
      replicate = i,
      h2total = h2total,
      pve_threshold = pve_threshold,
      n_causal = n_causal,
      Method = "SuSiE-ash (Scenario 6)",
      cs_fdr = metrics$cs_fdr,
      cs_recall = metrics$cs_recall,
      stringsAsFactors = FALSE
    )
    
    ## Method 4: Fineboost
    cs_fineboost <- if (!is.null(rep_data$fineboost_out$ucos_details$ucos$ucos_index))
      rep_data$fineboost_out$ucos_details$ucos$ucos_index else list()
    metrics <- compute_cs_metrics(cs_fineboost, truth)
    results_list_orig[[length(results_list_orig) + 1]] <- data.frame(
      file = filename,
      replicate = i,
      h2total = h2total,
      pve_threshold = pve_threshold,
      n_causal = n_causal,
      Method = "Fineboost",
      cs_fdr = metrics$cs_fdr,
      cs_recall = metrics$cs_recall,
      stringsAsFactors = FALSE
    )
  }
  rm(file_data)
  gc()
  message("Processed original file: ", filename)
}

results_orig_df <- do.call(rbind, results_list_orig)

# ------------------------------
# Process New Files (CS Metrics)
# ------------------------------

results_list_new <- list()

for (file in new_files) {
  filename <- basename(file)
  h2total <- as.numeric(str_extract(filename, "(?<=h2total)[0-9.]+"))
  Klength <- as.numeric(str_extract(filename, "(?<=Klength)[0-9.]+"))
  warmStart <- as.numeric(str_extract(filename, "(?<=warmStart)[0-9]+"))
  message("Processing new file: ", filename)
  file_data <- readRDS(file)
  
  for (i in seq_along(file_data$all_results)) {
    rep_data <- file_data$all_results[[i]]
    beta <- rep_data$data$beta
    residual_variance <- rep_data$data$residual_variance
    truth <- is_causal(beta, residual_variance, pve_threshold)
    
    for (grid in 1:6) {
      grid_label <- switch(as.character(grid),
                           "1" = "Grid 1: Base",
                           "2" = "Grid 2: 10x",
                           "3" = "Grid 3: 100x",
                           "4" = "Grid 4: 500x",
                           "5" = "Grid 5: 1000x",
                           "6" = "Grid 6: 10000x")
      grid_obj <- rep_data[[paste0("susie_ash_", grid)]]
      cs_susie_ash <- if (!is.null(grid_obj$sets)) grid_obj$sets else list()
      metrics <- compute_cs_metrics(cs_susie_ash, truth)
      method_label <- paste0("SuSiE.ash (", grid_label, "), K.length = ", Klength, ", Warmstart = ", warmStart)
      
      results_list_new[[length(results_list_new) + 1]] <- data.frame(
        file = filename,
        replicate = i,
        h2total = h2total,
        pve_threshold = pve_threshold,
        n_causal = n_causal,
        Method = method_label,
        cs_fdr = metrics$cs_fdr,
        cs_recall = metrics$cs_recall,
        stringsAsFactors = FALSE
      )
    }
  }
  rm(file_data)
  gc()
  message("Processed new file: ", filename)
}

results_new_df <- do.call(rbind, results_list_new)

# ------------------------------
# Combine Original and New CS Metrics
# ------------------------------

combined_df <- bind_rows(results_orig_df, results_new_df)

# Average Number of Causal Variants
avg_causal_df <- combined_df %>% 
  group_by(h2total) %>% 
  summarise(avg_num_causal = mean(n_causal, na.rm = TRUE))

# Summarize by Method and h2total for CS metrics
summary_by_h2 <- combined_df %>%
  group_by(Method, h2total) %>%
  summarise(
    mean_cs_fdr = mean(cs_fdr, na.rm = TRUE),
    sd_cs_fdr = sd(cs_fdr, na.rm = TRUE),
    se_cs_fdr = sd_cs_fdr / sqrt(n()),
    mean_cs_recall = mean(cs_recall, na.rm = TRUE),
    sd_cs_recall = sd(cs_recall, na.rm = TRUE),
    se_cs_recall = sd_cs_recall / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

# Filter out unwanted grids/methods and remove h2total == 0.3 (already filtered in file selection)
summary_by_h2_filtered <- summary_by_h2 %>%
  filter(!grepl("Grid 2: 10x|Grid 3: 100x|Grid 4: 500x|Grid 5: 1000x|Grid 6: 10000x|K.length = 10|K.length = 30|Warmstart = 0", Method)) %>%
  mutate(Method = case_when(
    Method == "SuSiE-ash (Scenario 6)" ~ "SuSiE.ash (Grid 2)",
    grepl("SuSiE\\.ash \\(Grid 1: Base\\)", Method) ~ "SuSiE.ash (Grid 1)",
    TRUE ~ Method
  )) %>% filter(Method != "SuSiE.ash (Grid 1)")

# Pivot the h2 summary for plotting CS metrics
plot_data_h2 <- summary_by_h2_filtered %>%
  pivot_longer(
    cols = c(mean_cs_fdr, mean_cs_recall, se_cs_fdr, se_cs_recall),
    names_to = c(".value", "Metric"),
    names_pattern = "(mean|se)_cs_(.*)"
  ) %>%
  mutate(Metric = ifelse(Metric == "fdr", "CS_FDR", "CS_Recall"))

# ------------------------------
# Plotting: CS FDR & Recall by h2total
# ------------------------------

# Function to create a 1x2 plot (side-by-side CS_FDR and CS_Recall) for a given heritability level
plot_by_h2 <- function(h2_val, data, pve_threshold, avg_causal) {
  data_h2 <- data %>% filter(h2total == h2_val)
  
  # CS_FDR plot with y-axis 0 to 0.75
  plot_fdr <- ggplot(data_h2 %>% filter(Metric == "CS_FDR"),
                     aes(x = Method, y = mean, fill = Method)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
    geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                  position = position_dodge(width = 0.7), width = 0.2) +
    labs(x = "Method",
         y = "CS FDR",
         title = paste("50 Replicates, h² =", h2_val),
         subtitle = paste0("Causal PVE threshold = ", pve_threshold, 
                          "\nAvg # Causal Variants = ", round(avg_causal, 2))) +
    scale_y_continuous(limits = c(0, 0.75)) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    geom_hline(yintercept = 0.05, color = "red", linetype = "dotted")
  
  # CS_Recall plot with y-axis 0 to 1
  plot_recall <- ggplot(data_h2 %>% filter(Metric == "CS_Recall"),
                        aes(x = Method, y = mean, fill = Method)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
    geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                  position = position_dodge(width = 0.7), width = 0.2) +
    labs(x = "Method",
         y = "CS Recall",
         title = paste("50 Replicates, h² =", h2_val),
         subtitle = paste0("Causal PVE threshold = ", pve_threshold, 
                          "\nAvg # Causal Variants = ", round(avg_causal, 2))) +
    scale_y_continuous(limits = c(0, 1)) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none")
  
  combined_plot <- plot_fdr + plot_recall
  return(combined_plot)
}


# Example calls for CS metrics plots (for h² = 0.15 and 0.50)
avg_015 <- avg_causal_df %>% filter(h2total == 0.15) %>% pull(avg_num_causal)
avg_050 <- avg_causal_df %>% filter(h2total == 0.50) %>% pull(avg_num_causal)

cs_plot_015 <- plot_by_h2(0.15, plot_data_h2, pve_threshold, avg_015)
cs_plot_050 <- plot_by_h2(0.50, plot_data_h2, pve_threshold, avg_050)

print(cs_plot_015)
print(cs_plot_050)

```



```{r, echo=FALSE, warning=F, message=F}
# ---------------------------------------------------------------------------------
# Now, Compute AUROC and AUPRC using Top N variants (by absolute effect size)
# ---------------------------------------------------------------------------------
# 
# results_list_auroc <- list()
# results_list_auprc <- list()
# 
# # --- Process Original Files for ROC metrics ---
# for (file in orig_files) {
#   filename <- basename(file)
#   h2total <- as.numeric(str_extract(filename, "(?<=h2total)[0-9.]+"))
#   message("Processing original file for ROC metrics: ", filename)
#   file_data <- readRDS(file)
#   
#   for (i in seq_along(file_data$all_results)) {
#     rep_data <- file_data$all_results[[i]]
#     betas <- rep_data$data$beta
#     p <- length(betas)
#     ranked_indices <- order(abs(betas), decreasing = TRUE)
#     
#     for (n in top_N) {
#       truth <- rep(0, p)
#       truth[ranked_indices[1:n]] <- 1
#       
#       ## Method 1: SuSiE
#       pips <- rep_data$susie_out$pip
#       auroc_val <- compute_auroc(truth, pips)
#       auprc_val <- compute_auprc(truth, pips)
#       results_list_auroc[[length(results_list_auroc) + 1]] <- data.frame(
#         file = filename, replicate = i, h2total = h2total, top_N = n,
#         Method = "SuSiE", AUROC = auroc_val, stringsAsFactors = FALSE
#       )
#       results_list_auprc[[length(results_list_auprc) + 1]] <- data.frame(
#         file = filename, replicate = i, h2total = h2total, top_N = n,
#         Method = "SuSiE", AUPRC = auprc_val, stringsAsFactors = FALSE
#       )
#       
#       ## Method 2: SuSiE-inf
#       pips <- rep_data$susie_inf_out$marginal_PIP
#       auroc_val <- compute_auroc(truth, pips)
#       auprc_val <- compute_auprc(truth, pips)
#       results_list_auroc[[length(results_list_auroc) + 1]] <- data.frame(
#         file = filename, replicate = i, h2total = h2total, top_N = n,
#         Method = "SuSiE-inf", AUROC = auroc_val, stringsAsFactors = FALSE
#       )
#       results_list_auprc[[length(results_list_auprc) + 1]] <- data.frame(
#         file = filename, replicate = i, h2total = h2total, top_N = n,
#         Method = "SuSiE-inf", AUPRC = auprc_val, stringsAsFactors = FALSE
#       )
#       
#       ## Method 3: SuSiE-ash (Scenario 6)
#       pips <- rep_data$susie_ash_6$marginal_PIP
#       auroc_val <- compute_auroc(truth, pips)
#       auprc_val <- compute_auprc(truth, pips)
#       results_list_auroc[[length(results_list_auroc) + 1]] <- data.frame(
#         file = filename, replicate = i, h2total = h2total, top_N = n,
#         Method = "SuSiE-ash (Scenario 6)", AUROC = auroc_val, stringsAsFactors = FALSE
#       )
#       results_list_auprc[[length(results_list_auprc) + 1]] <- data.frame(
#         file = filename, replicate = i, h2total = h2total, top_N = n,
#         Method = "SuSiE-ash (Scenario 6)", AUPRC = auprc_val, stringsAsFactors = FALSE
#       )
#       
#       ## Method 4: Fineboost
#       pips <- rep_data$fineboost_out$pip
#       auroc_val <- compute_auroc(truth, pips)
#       auprc_val <- compute_auprc(truth, pips)
#       results_list_auroc[[length(results_list_auroc) + 1]] <- data.frame(
#         file = filename, replicate = i, h2total = h2total, top_N = n,
#         Method = "Fineboost", AUROC = auroc_val, stringsAsFactors = FALSE
#       )
#       results_list_auprc[[length(results_list_auprc) + 1]] <- data.frame(
#         file = filename, replicate = i, h2total = h2total, top_N = n,
#         Method = "Fineboost", AUPRC = auprc_val, stringsAsFactors = FALSE
#       )
#     }
#   }
#   rm(file_data)
#   gc()
#   message("Processed original file for ROC metrics: ", filename)
# }
# 
# # --- Process New Files for ROC metrics ---
# for (file in new_files) {
#   filename <- basename(file)
#   h2total <- as.numeric(str_extract(filename, "(?<=h2total)[0-9.]+"))
#   Klength <- as.numeric(str_extract(filename, "(?<=Klength)[0-9.]+"))
#   warmStart <- as.numeric(str_extract(filename, "(?<=warmStart)[0-9]+"))
#   message("Processing new file for ROC metrics: ", filename)
#   file_data <- readRDS(file)
#   
#   for (i in seq_along(file_data$all_results)) {
#     rep_data <- file_data$all_results[[i]]
#     betas <- rep_data$data$beta
#     p <- length(betas)
#     ranked_indices <- order(abs(betas), decreasing = TRUE)
#     
#     for (n in top_N) {
#       truth <- rep(0, p)
#       truth[ranked_indices[1:n]] <- 1
#       
#       for (grid in 1:6) {
#         grid_label <- switch(as.character(grid),
#                              "1" = "Grid 1: Base",
#                              "2" = "Grid 2: 10x",
#                              "3" = "Grid 3: 100x",
#                              "4" = "Grid 4: 500x",
#                              "5" = "Grid 5: 1000x",
#                              "6" = "Grid 6: 10000x")
#         grid_obj <- rep_data[[paste0("susie_ash_", grid)]]
#         pips <- grid_obj$marginal_PIP
#         auroc_val <- compute_auroc(truth, pips)
#         auprc_val <- compute_auprc(truth, pips)
#         
#         method_label <- paste0("SuSiE.ash (", grid_label, "), K.length = ", Klength, ", Warmstart = ", warmStart)
#         
#         results_list_auroc[[length(results_list_auroc) + 1]] <- data.frame(
#           file = filename, replicate = i, h2total = h2total, top_N = n,
#           Method = method_label, AUROC = auroc_val, stringsAsFactors = FALSE
#         )
#         results_list_auprc[[length(results_list_auprc) + 1]] <- data.frame(
#           file = filename, replicate = i, h2total = h2total, top_N = n,
#           Method = method_label, AUPRC = auprc_val, stringsAsFactors = FALSE
#         )
#       }
#     }
#   }
#   rm(file_data)
#   gc()
#   message("Processed new file for ROC metrics: ", filename)
# }
# 
# results_auroc_df <- do.call(rbind, results_list_auroc)
# results_auprc_df <- do.call(rbind, results_list_auprc)
# 
# # ------------------------------
# # Summarize and Filter AUROC/AUPRC Results (Stratify by h2total = 0.15 and 0.50)
# # ------------------------------
# 
# summary_auroc_df <- results_auroc_df %>%
#   filter(h2total %in% c(0.15, 0.50)) %>%
#   filter(!grepl("Grid 2: 10x|Grid 3: 100x|Grid 4: 500x|Grid 5: 1000x|Grid 6: 10000x|K.length = 10|K.length = 30|Warmstart = 0", Method)) %>%
#   mutate(Method = case_when(
#     Method == "SuSiE-ash (Scenario 6)" ~ "SuSiE.ash (Grid 2)",
#     grepl("SuSiE\\.ash \\(Grid 1: Base\\)", Method) ~ "SuSiE.ash (Grid 1)",
#     TRUE ~ Method
#   )) %>%
#   group_by(top_N, Method, h2total) %>%
#   summarise(
#     mean_AUROC = mean(AUROC, na.rm = TRUE),
#     sd_AUROC = sd(AUROC, na.rm = TRUE),
#     se_AUROC = sd_AUROC / sqrt(n()),
#     n = n(),
#     .groups = "drop"
#   )
# 
# summary_auprc_df <- results_auprc_df %>%
#   filter(h2total %in% c(0.15, 0.50)) %>%
#   filter(!grepl("Grid 2: 10x|Grid 3: 100x|Grid 4: 500x|Grid 5: 1000x|Grid 6: 10000x|K.length = 10|K.length = 30|Warmstart = 0", Method)) %>%
#   mutate(Method = case_when(
#     Method == "SuSiE-ash (Scenario 6)" ~ "SuSiE.ash (Grid 2)",
#     grepl("SuSiE\\.ash \\(Grid 1: Base\\)", Method) ~ "SuSiE.ash (Grid 1)",
#     TRUE ~ Method
#   )) %>%
#   group_by(top_N, Method, h2total) %>%
#   summarise(
#     mean_AUPRC = mean(AUPRC, na.rm = TRUE),
#     sd_AUPRC = sd(AUPRC, na.rm = TRUE),
#     se_AUPRC = sd_AUPRC / sqrt(n()),
#     n = n(),
#     .groups = "drop"
#   )
# 
# # ------------------------------
# # Plotting: AUROC and AUPRC by h2total
# # ------------------------------
# 
# # AUROC Plot Function (for a given h2total)
# plot_auroc_by_h2 <- function(h2_val, data) {
#   data %>% filter(h2total == h2_val) %>%
#     ggplot(aes(x = factor(top_N), y = mean_AUROC, color = Method, group = Method)) +
#     geom_point(position = position_dodge(width = 0.5), size = 3) +
#     geom_errorbar(aes(ymin = mean_AUROC - se_AUROC, ymax = mean_AUROC + se_AUROC),
#                   position = position_dodge(width = 0.5), width = 0.2) +
#     labs(x = "Number of Causal Variants (Top N)",
#          y = "Mean AUROC",
#          title = paste("AUROC, h² =", h2_val)) +
#     theme_minimal()
# }
# 
# # AUPRC Plot Function (for a given h2total)
# plot_auprc_by_h2 <- function(h2_val, data) {
#   data %>% filter(h2total == h2_val) %>%
#     ggplot(aes(x = factor(top_N), y = mean_AUPRC, color = Method, group = Method)) +
#     geom_point(position = position_dodge(width = 0.5), size = 3) +
#     geom_errorbar(aes(ymin = mean_AUPRC - se_AUPRC, ymax = mean_AUPRC + se_AUPRC),
#                   position = position_dodge(width = 0.5), width = 0.2) +
#     labs(x = "Number of Causal Variants (Top N)",
#          y = "Mean AUPRC",
#          title = paste("AUPRC, h² =", h2_val)) +
#     theme_minimal()
# }
# 
# # Create AUROC and AUPRC plots for h2total = 0.15 and 0.50
# auroc_plot_015 <- plot_auroc_by_h2(0.15, summary_auroc_df)
# auroc_plot_050 <- plot_auroc_by_h2(0.50, summary_auroc_df)
# auprc_plot_015 <- plot_auprc_by_h2(0.15, summary_auprc_df)
# auprc_plot_050 <- plot_auprc_by_h2(0.50, summary_auprc_df)
# 
# # Optionally, arrange the plots:
# # grid.arrange(auroc_plot_015, auroc_plot_050, ncol = 1)
# # grid.arrange(auprc_plot_015, auprc_plot_050, ncol = 1)
# 
# # Or print them individually:
# print(auroc_plot_015)
# print(auroc_plot_050)
# print(auprc_plot_015)
# print(auprc_plot_050)
```


