---
title: "Rcpp Comparison"
output: html_document
date: "2024-08-29"
---

```{r, include = F}
knitr::opts_knit$set(root.dir = "/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH")
library(Rcpp)
library(RcppArmadillo)
library(microbenchmark)
library(dplyr)

sourceCpp("code/rcpp-testing.cpp")
```

```{r, echo = F}
# Function to generate data
generate_data <- function() {
  X <- readRDS("/Users/alexmccreight/Columbia/data/X4")
  X <- scale(X, center = TRUE, scale = TRUE)
  n <- nrow(X)
  p <- ncol(X)
  MAF <- 0.1
  Ltrue <- 5
  ssq <- 0.01
  sigmasq <- 1
  b <- rep(0, p)
  inds <- sample(p, Ltrue, replace = FALSE)
  b[inds] <- rnorm(Ltrue) * sqrt(ssq)
  tausq <- 5e-4
  theta <- rnorm(p) * sqrt(tausq)
  effects <- X %*% b + X %*% theta
  y <- effects + rnorm(n) * sqrt(sigmasq)

  # Compute additional values needed for comparison functions
  XtX_val <- t(X) %*% X
  LD <- XtX_val / n
  eig <- eigen(LD, symmetric = TRUE)
  V <- eig$vectors[, ncol(eig$vectors):1]
  Dsq <- pmax(n * sort(eig$values), 0)
  z <- (t(X) %*% y) / sqrt(n)
  Xty <- sqrt(n) * z
  var <- tausq * Dsq + sigmasq
  VtXty <- t(V) %*% Xty

  list(X=X, y=y, LD=LD, V=V, Dsq=Dsq, Xty=Xty, var=var, VtXty=VtXty)
}

# Function to compare R and Rcpp versions
compare_functions <- function(r_func, rcpp_func, ..., times = 10) {
  benchmark <- microbenchmark(
    R = r_func(...),
    Rcpp = rcpp_func(...),
    times = times
  )
  r_time <- mean(benchmark$time[benchmark$expr == "R"]) / 1e9  # Convert to seconds
  rcpp_time <- mean(benchmark$time[benchmark$expr == "Rcpp"]) / 1e9  # Convert to seconds
  percent_change <- (rcpp_time - r_time) / r_time * 100
  c(r_time = r_time, rcpp_time = rcpp_time, percent_change = percent_change)
}
```

# Rcpp functions versus R functions

Here we compare the computation speed of the following objects: `XtX`, `Xty`, `Eigendecomposition`, `VtXty`, and `diagXtOmegaX`. For each iteration of this comparison we will first simulate data, compute all of the previously mentioned objects 10 times each and average their computation times for both Rcpp and R. We then repeat this process for a total of 10 iterations and take the average yielding three key outputs: `Average R Time`, `Average Rcpp Time`, and `Percent Change`, which measures relative difference in average computation time between Rcpp and R functions. A negative percentage change indicates a decrease in computation time, while a positive change indicates an increase computation time.

```{r, echo = F, warning = F}
# Number of iterations
num_iterations <- 10

# Initialize results dataframe
results <- data.frame()

for (i in 1:num_iterations) {
  # Generate new data for each iteration
  data <- generate_data()

  # Run comparisons
  comparisons <- list(
    XtX = compare_functions(function(X) t(X) %*% X, compute_XtX, data$X),
    Xty = compare_functions(function(X, y) t(X) %*% y, compute_Xty, data$X, data$y),
    eigen_sym = compare_functions(function(LD) eigen(LD, symmetric = TRUE), compute_eigen_sym, data$LD),
    VtXty = compare_functions(function(V, Xty) t(V) %*% Xty, compute_VtXty, data$V, data$Xty),
    diagXtOmegaX = compare_functions(
      function(V, Dsq, var) rowSums(sweep(V^2, 2, (Dsq / var), `*`)),
      compute_diagXtOmegaX, data$V, data$Dsq, data$var
    ),
    XtOmegay = compare_functions(
      function(V, VtXty, var) V %*% (VtXty / var),
      compute_XtOmegay, data$V, data$VtXty, data$var
    )
  )

  # Add results to dataframe
  iteration_results <- do.call(rbind, comparisons) %>%
    as.data.frame() %>%
    tibble::rownames_to_column("Function") %>%
    mutate(Iteration = i)

  results <- rbind(results, iteration_results)
}

# Calculate average results across all iterations
average_results <- results %>%
  group_by(Function) %>%
  summarize(
    Avg_R_Time = mean(r_time),
    Avg_Rcpp_Time = mean(rcpp_time),
    Avg_Percent_Change = mean(percent_change)
  )

# Print average results
print(average_results)
```

