---
title: "SuSiE-ASH Simulation"
output: html_document
date: "2024-03-20"
---

Note-to-self: 
- Recall/Power = Percentage of simulated LARGE EFFECTS among the top N variants when ranked by PIP (Note that recall for SuSiE-Inf was very similar to, but slightly lower than the recall of SuSiE)
- False Discovery Rate = P(Non-Causal | PIP > 0.9)
- Calibration = Of variants with PIP = x%, we expect x% are truly causal
- Replication Failure Rate = P(PIP Large Sample < 0.1 | PIP Small Sample > 0.9)
- Expected Proportion of Non-Causal Variants (EPN) --> Supp table 6,7,8
- Coverage: proportion of variants with nonzero effects
- Heritability: proportion of outcome variance that can be attributed to genetic effects (our betas and thetas)


If the RFR is significantly higher than the EPN, it suggests miscalibration.

```{r, include = F}
# Load Libraries
library(tidyverse)
library(mr.ash.alpha)
library(susieR)
library(FamilyRank)

# Fix File path later
devtools::load_all('/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/submodules/susieR')
devtools::load_all('/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/submodules/mr.ash.alpha')

#devtools::install_github("stephenslab/mr.ash.alpha")

# SuSiE-ASH (v1; remaining effects)
source("/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/code/SuSiE-ASH-Versions/susie-ash-warmstart-v1.R")

# SuSiE-ASH (v2; Gao's Version)
source("/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/code/SuSiE-ASH-Versions/susie-ash-warmstart-v2.R")
```

# Simulated Data

```{r}
generate_data <- function(n, p, heritability, sparse_coverage, nonsparse_coverage) {
  # Generate sparse effects (beta.true)
  num_sparse <- round(p * sparse_coverage)
  beta.true <- rep(0, p)
  #beta.true[sample(p, num_sparse)] <- rnorm(num_sparse, 0, 0.5)
  beta.true[sample(p, num_sparse)] <- rbinorm(num_sparse, mean1 = 0.4, mean2 = -0.4, sd1 = 0.15, sd2 = 0.2, prop = 0.5)
  
  # Generate nonsparse effects (theta.true)
  num_nonsparse <- round(p * nonsparse_coverage)
  theta.true <- rep(0, p)
  
  # Set the mixture prior parameters for nonsparse effects
  nonsparse_pis <- c(0.5, 0.3, 0.15, 0.05)
  #nonsparse_sds <- c(0.0001, 0.002, 0.05, 0.1) # William Mixture
  #nonsparse_sds <- c(0.0001, 0.005, 0.0075, 0.05) 
  nonsparse_sds <- c(0.0001, 0.002/2, 0.05/2, 0.1/2)
  
  theta.true[sample(p, num_nonsparse)] <- sapply(1:num_nonsparse, function(i) {
    dist.sel <- runif(1)
    
    if (dist.sel < nonsparse_pis[1]) {
      val <- 0
    } else if (dist.sel < sum(nonsparse_pis[1:2])) {
      val <- rnorm(1, 0, nonsparse_sds[2])
    } else if (dist.sel < sum(nonsparse_pis[1:3])) {
      val <- rnorm(1, 0, nonsparse_sds[3]) 
    } else {
      val <- rnorm(1, 0, nonsparse_sds[4])
    }
    
    val
  })
  
  # Create Design Matrix
  X <- cbind(matrix(rnorm(n*p),nrow=n))
  X <- scale(X, center=TRUE, scale=FALSE)

  # Create Residual Error
  sigmasq_error <- (var(X %*% beta.true)*(1 - heritability)) / heritability

  # Create Outcomes
  y <- X%*%matrix(beta.true,ncol=1) + X%*%matrix(theta.true,ncol=1) + rnorm(n,0,sqrt(sigmasq_error))
  y <- scale(y, center=TRUE, scale=FALSE)
  
  # Store Information
  return(list(X = X, y = y, error = sigmasq_error, beta = beta.true, theta = theta.true))
}

data <- generate_data(n = 10000, p = 500, heritability = 0.75, sparse_coverage = 0.01, nonsparse_coverage = 0.1)

plot(data$beta, ylim = c(-1.5,1.5), ylab = "beta")
plot(data$theta, ylim = c(-.5,.5), ylab = "theta")
```

```{r}
output <- susie_ash_warmstart(X = data$X, y = data$y, L = 10, tol = 0.02)
output2 <- susie(X = data$X, y = data$y, L = 10, tol = 0.001)
output3 <- susie_ash_warmstart_re(X = data$X, y = data$y, L = 10, tol = 0.02)

susie_get_cs(output, X = data$X)


beta.true.index = which(data$beta != 0)
theta.true.index = which(abs(data$theta) >= 0.025)

# Color + Shape Differ for True Effects
color_vector <- rep("black", length(output$pip))
color_vector[beta.true.index] <- "red"
shape_vector <- rep(1, length(output$pip))
shape_vector[beta.true.index] <- 17

color_vector[theta.true.index] <- "blue"
shape_vector[theta.true.index] <- 16

output$pip %>% plot(., main = "SuSiE-ASH (v2) PIP", col = color_vector, pch = shape_vector) %>% abline(h = 0.9, col="black", lty = 2) 
output2$pip %>% plot(., main = "SuSiE PIP", col = color_vector, pch = shape_vector) %>% abline(h = 0.9, col="black", lty = 2) 
output3$pip %>% plot(., main = "SuSiE-ASH (v1) PIP", col = color_vector, pch = shape_vector) %>% abline(h = 0.9, col="black", lty = 2) 

```



# Methods (SuSiE, SuSiE-ASH Variants)

```{r}
method_and_score <- function(X = data$X, y = data$y, beta = data$beta, L = 10, threshold = 0.9) {
  # Run various methods
  susie_output <- susie(X = X, y = y, L = L)
  susie_ash_output_v1 <- susie_ash_warmstart_re(X = X, y = y, L = L, warm_start = 5, tol = 0.03)
  susie_ash_output_v2 <- susie_ash_warmstart(X = X, y = y, L = L, warm_start = 5, tol = 0.03)
  
  
  calc_metrics <- function(mod, beta = data$beta, threshold = 0.9) {
    # Identify causal variables (non-zero sparse effects)
    causal <- beta != 0
    significant <- mod$pip > threshold
    
    # Calculate FDR
    fdr <- ifelse(sum(significant) > 0, sum(!causal & significant) / sum(significant), 0)
    
    # Calculate Recall
    recall <- ifelse(sum(causal) > 0, sum(causal & significant) / sum(causal), 0)
    
    # Calculate Average CS Size
    cs_size <- length(unlist(susie_get_cs(mod, X = data$X, coverage = threshold)$cs)) / length(susie_get_cs(mod, X = data$X, coverage = threshold)$cs)
    
    # Calculate Coverage (proportion of cs with causal effect)
    coverage_tmp <- sum(which(causal) %in% susie_get_cs(output, X = data$X, coverage = threshold)$cs) / length(susie_get_cs(mod, X = data$X, coverage = threshold)$cs)
      
    
    # CS Based FDR and Recall
    # test.true <- which(data$X != 0)
    # test.cs <- susie_get_cs(mod, X = data$X)$cs
    # 
    #   TP = sum(test.true %in% unlist(test.cs))
    #   FN = length(test.true) - TP
    #   FP = length(test.cs) - lapply(1:length(test.cs), function(cs.l){ ifelse(sum(test.cs[[cs.l]] %in%test.true)!=0,T,F)}) %>% unlist(.) %>% sum(.) # FP based on CSs
    #   FP = length(unlist(test.cs)) - sum(unlist(test.cs) %in% test.true) # FP based on elements
    #   
    #   cs_fdr = FP/(TP+FP)
    #   cs_recall = TP/(TP+FN)

    
    return(list(fdr = fdr, recall = recall, cs_size = cs_size, coverage_tmp = coverage_tmp))#, cs_fdr = cs_fdr, cs_recall = cs_recall))
  }
  
  # Calculate FDR and Recall for each method
  susie_metrics <- calc_metrics(susie_output, beta, threshold)
  susie_ash_v1_metrics <- calc_metrics(susie_ash_output_v1, beta, threshold)
  susie_ash_v2_metrics <- calc_metrics(susie_ash_output_v2, beta, threshold)
  
  #Create a data frame with the results
  metrics_table  <- data.frame(
    Model = c("SuSiE", "SuSiE-ASH v1", "SuSiE-ASH v2"),
    FDR = c(susie_metrics$fdr, susie_ash_v1_metrics$fdr, susie_ash_v2_metrics$fdr),
    Recall = c(susie_metrics$recall, susie_ash_v1_metrics$recall, susie_ash_v2_metrics$recall),
    CS_Size = c(susie_metrics$cs_size, susie_ash_v1_metrics$cs_size, susie_ash_v2_metrics$cs_size),
    Coverage_tmp = c(susie_metrics$coverage_tmp, susie_ash_v1_metrics$coverage_tmp, susie_ash_v2_metrics$coverage_tmp)
    #CS_FDR = c(susie_metrics$cs_fdr, susie_ash_v1_metrics$cs_fdr, susie_ash_v2_metrics$cs_fdr),
    #CS_Recall = c(susie_metrics$cs_recall, susie_ash_v1_metrics$cs_recall, susie_ash_v2_metrics$cs_recall)
  )
  
  # metrics_table  <- data.frame(
  #   Model = c("SuSiE", "SuSiE-ASH"),
  #   FDR = c(susie_metrics$fdr, susie_ash_v2_metrics$fdr),
  #   Recall = c(susie_metrics$recall, susie_ash_v2_metrics$recall)
  # )
  
  # Return the results table
  return(list(
    metrics = metrics_table,
    susie_output = susie_output,
    susie_ash_output_v1 = susie_ash_output_v1,
    susie_ash_output_v2 = susie_ash_output_v2,
    betas = data$beta,
    thetas = data$theta)
  )
}

#tmp <- method_and_score()
```


# Scores (FDR, Recall)

```{r}
simulate_and_score <- function(num_simulations = 10, n = 10000, p = 500, heritability = 0.75, sparse_coverage = 0.01, nonsparse_coverage = 0.1, L = 10, threshold = 0.9) {
  
  # Initialize lists to store results
  all_metrics <- list()
  all_betas <- list()
  all_thetas <- list()
  all_susie_outputs <- list()
  all_susie_ash_outputs_v1 <- list()
  all_susie_ash_outputs_v2 <- list()
  all_seeds <- numeric(num_simulations)
  
  for (i in 1:num_simulations) {
    cat("Running simulation", i, "out of", num_simulations, "\n")
    
    # Set random seed for each simulation
    seed <- abs(round(rnorm(1, mean = 0, sd = 1000)))
    set.seed(seed)
    
    # Generate data
    data <- generate_data(n, p, heritability, sparse_coverage, nonsparse_coverage)
    
    # Run methods and calculate metrics
    results <- method_and_score(data$X, data$y, data$beta, L, threshold)
    
    # Store results + betas/thetas
    all_metrics[[i]] <- results$metrics
    all_betas[[i]] <- data$beta
    all_thetas[[i]] <- data$theta
    all_susie_outputs[[i]] <- results$susie_output
    all_susie_ash_outputs_v1[[i]] <- results$susie_ash_output_v1
    all_susie_ash_outputs_v2[[i]] <- results$susie_ash_output_v2
    all_seeds[i] <- seed
  }
  
  # Calculate average metrics
  avg_metrics <- data.frame(
    Model = unique(all_metrics[[1]]$Model),
    FDR = Reduce("+", lapply(all_metrics, function(x) x$FDR)) / num_simulations,
    Recall = Reduce("+", lapply(all_metrics, function(x) x$Recall)) / num_simulations,
    CS_Size = Reduce("+", lapply(all_metrics, function(x) x$CS_Size)) / num_simulations,
    Coverage_tmp = Reduce("+", lapply(all_metrics, function(x) x$Coverage_tmp)) / num_simulations
    #CS_FDR = Reduce("+", lapply(all_metrics, function(x) x$CS_FDR)) / num_simulations,
    #CS_Recall = Reduce("+", lapply(all_metrics, function(x) x$CS_Recall)) / num_simulations
  )
  
  # Return all results
  return(list(
    avg_metrics = avg_metrics,
    all_metrics = all_metrics,
    all_betas = all_betas,
    all_thetas = all_thetas,
    all_susie_outputs = all_susie_outputs,
    all_susie_ash_outputs_v1 = all_susie_ash_outputs_v1,
    all_susie_ash_outputs_v2 = all_susie_ash_outputs_v2,
    all_seeds = all_seeds
  ))
}

large_simulation <- simulate_and_score(num_simulations = 2, n = 10000, p = 500, heritability = 0.75, sparse_coverage = 0.01, nonsparse_coverage = 0.1, L = 10, threshold = 0.9)

print(large_simulation$avg_metrics)
print(large_simulation$all_seeds)


```

Note-to-self: 
- Recall/Power = Percentage of simulated LARGE EFFECTS among the top N variants when ranked by PIP (Note that recall for SuSiE-Inf was very similar to, but slightly lower than the recall of SuSiE)
- False Discovery Rate = P(Non-Causal | PIP > 0.9)
- Calibration = Of variants with PIP = x%, we expect x% are truly causal
- Replication Failure Rate = P(PIP Large Sample < 0.1 | PIP Small Sample > 0.9)
- Expected Proportion of Non-Causal Variants (EPN) --> Supp table 6,7,8
- Coverage: proportion of variants with nonzero effects
- Heritability: proportion of outcome variance that can be attributed to genetic effects (our betas and thetas)


To implement: 
1) CS Size, similar to precision

2) Coverage = proportion of credible sets with a causal effect, 95% coverage want CS with 95% prob of containing a causal effect

3) ROC curve with FDR vs Power, way of comparing PIP

4) Coverage versus L (from susie) graph 

5) How does coverage change when we vary heritability versus the amount of actual simulated effects
