---
title: "Simulation Tables"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH")
library(tidyverse)
```

# Algorithms

| Algorithm           | Description                                                                 |
|:--------------------|:----------------------------------------------------------------------------|
| **SuSiE**           | SuSiE algorithm.                                                            |
| **mr.ash**          | mr.ash algorithm.                                                           |
| **SuSiE-ash**       | Original SuSiE-ash algorithm. (algorithm 1 on Overleaf)                                               |
| **SuSiE-ash (v4)**  | Run SuSiE until convergence, then run mr.ash on residuals.                  |
| **SuSiE-ash (v5)**  | Run mr.ash until convergence, then run SuSiE on residuals.                  |
| **SuSiE-ash (v10)** | New SuSiE-ash algorithm (algorithm 2 on Overleaf)                                                    |
| **SuSiE-ash (v11)** | New SuSiE-ash algorithm (except update on all effects, not just high heritability ones). |
| **SuSiE-inf**       | SuSiE-inf algorithm.  

# Simulation Parameters

| Parameter                    | Description                                                                 |
|:-----------------------------|:----------------------------------------------------------------------------|
| **Heritability**             | Amount of variation in outcome explained by predictors.                                      |
| **Nonsparse**                | Proportion of infinitesimal effects with non-zero effects on outcome.                                         |
| **Theta to Beta Ratio**      | Ratio of heritability explained by nonsparse effects versus sparse effects.                                |

# Metrics

| Metric        | Description                                                                                      |
|:--------------|:-------------------------------------------------------------------------------------------------|
| **RMSE_y**    | Root Mean Squared Error of the outcome variable.                                                  |
| **CS_FDR**    | Credible set based FDR. |
| **CS_Recall** | Credible set based Recall. |
| **CS_Size**   | Average number of elements in the credible sets.                    |
| **Coverage**  | Proportion of credible sets with at least one causal variant.                       |



# Tables

```{r, echo = F, message = F}
# Read in results
tables <- read_csv("simulation/simulation avg metrics/avg_metrics_results_modified.csv")

# Split tables by simulation settings
split_tables <- tables %>%
  group_by(totHeritability, sparseEffect, nonsparse, ratio) %>%
  group_split()

for (i in seq_along(split_tables)) {
  simulation <- split_tables[[i]]
  settings <- simulation %>%
    select(totHeritability, sparseEffect, nonsparse, ratio) %>%
    distinct() %>%
    unlist() %>%
    as.list()
  
  cat("Simulation Setting", i, ": Heritability =", settings$totHeritability, 
      ", Nonsparse =", settings$nonsparse, 
      ", Theta to Beta Ratio =", settings$ratio, "\n")
  
  # Drop the simulation parameter columns
  simulation <- simulation %>%
    select(-totHeritability, -sparseEffect, -nonsparse, -ratio, -RMSE_beta, -RMSE_theta)
  
  print(simulation)
  cat("\n")
}
```

