---
title: "Synthetic Association"
output: html_document
---

The following vignette will explore a replicate where many non-causal variants are in near-perfect to moderate LD with two other causal variants creating a synthetic association. We will compare the difference in model output between SuSiE and Fineboost. 

# Moderate LD (|r| < 0.5) Sparse Simulation Setting

```{r, message = F, warning = F}
library(tidyverse)
LD_mat <- readRDS("/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/seed9_ld.rds")
data <- readRDS("/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/seed9_data.rds")
result <- readRDS("/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/seed9_res.rds")
```

# Causal Indices, Effect Sizes, and LD

This replicate contains 5 causal variants where the two strongest ones (ID = 408 and 565) are in moderate LD (|r| = 0.43).

```{r}
data$causal_indices
data$beta[data$causal_indices] %>% round(.,2)

LD_mat[data$causal_indices, data$causal_indices] %>% round(.,2) %>% abs()
```

## SuSiE vs Fineboost Credible Sets

Both SuSiE and Fineboost capture a single CS of size 70 and 77 respectively. Fineboost managed to capture all 70 variants found in the SuSiE CS and added 7 additional variants INCLUDING a causal variant. 

```{r, echo =F}
cat("SuSiE CS:\n")
susie_cs <- result$susie_out$sets$cs$L1; susie_cs

cat("\n Fineboost CS:\n")
fineboost_cs <- result$fineboost_out$ucos_details$ucos$ucos_index$`ucos1:y1`; sort(fineboost_cs)

#intersect(fineboost_cs, susie_cs) %>% length()
```

## LD Between Credible SuSiE's Credible Set and Causal Variants

All variants in SuSiE's credible set are in near perfect LD with a causal variant and moderate LD with another. 

```{r}
LD_mat[susie_cs, data$causal_indices] %>% abs() %>% round(.,2)
#LD_mat[sort(fineboost_cs), data$causal_indices] %>% abs() %>% round(.,2)
```

