---
title: "Failure of the mr.ash component to update after the first iteration of SuSiE"
output:
  html_document:
    code_folding: hide
---

```{r, include = F}
# Libraries
library(tidyverse)
library(Rcpp)
library(RcppArmadillo)

# Packages
source("/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/code/susie_versions/SuSiE_Ash_Marginal.R")
devtools::load_all("/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/new-rcpp/mr.ash.alpha", recompile = T)
devtools::load_all('/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/submodules/susieR')

# Data
precomputed_matrices <- readRDS("/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/vignettes/precomputed_matrices.rds")
X20 <- readRDS("/Users/alexmccreight/Columbia/data/X20")
X_subset <- readRDS("/Users/alexmccreight/Columbia/Research/SuSiE-ASH/SuSiE-ASH/X_subset")

# Data Generation Function
generate_eqtl_data <- function(X,
                               h2_total = 0.3,            # Total heritability. h2
                               prop_h2_sparse = 0.65,     # Proportion of h2_total explained by sparse effects (including sentinel).
                               prop_h2_oligogenic = 0.20, # Proportion of h2_total explained by oligogenic effects
                               prop_h2_infinitesimal = 0.15, # Proportion of h2_total explained by infinitesimal effects
                               prop_h2_sentinel = 0.7,    # Proportion of h2_sparse explained by sentinel SNP. h2_s
                               n_oligogenic = 20,
                               mixture_props = c(0.6, 0.4), # Adjusted mixture proportions
                               mixture_sds = c(0.0025, 0.005), # Standard deviations for mixture components
                               seed = NULL) {
  if (!is.null(seed)) set.seed(seed)

  ori.X <- X
  X <- scale(X)

  n_samples <- nrow(X)
  n_features <- ncol(X)

  # Calculate effect sizes for each component
  h2_sparse <- h2_total * prop_h2_sparse #h2_cojo --> change to 30%?
  h2_sentinel <- h2_sparse * prop_h2_sentinel
  h2_other_sparse <- h2_sparse - h2_sentinel
  h2_oligogenic <- h2_total * prop_h2_oligogenic
  h2_infinitesimal <- h2_total * prop_h2_infinitesimal

  # Generate effect sizes
  beta <- rep(0, n_features)

  # Sentinel SNP effect (largest effect among sparse effects)
  sentinel_index <- sample(1:n_features, 1)
  beta[sentinel_index] <- rnorm(1, 0, sqrt(h2_sentinel))

  # Other sparse effects (large and mappable)
  n_other_sparse <- 2
  other_sparse_indices <- sample(setdiff(1:n_features, sentinel_index), n_other_sparse)
  if (n_other_sparse > 0) {
    # Distribute h2_other_sparse equally among other sparse SNPs
    beta[other_sparse_indices] <- rnorm(n_other_sparse, 0, sqrt(h2_other_sparse / n_other_sparse))
    # Ensure the sentinel SNP has the largest effect size
    max_other_sparse_effect <- max(abs(beta[other_sparse_indices]))
    if (abs(beta[sentinel_index]) <= max_other_sparse_effect) {
      beta[sentinel_index] <- sign(beta[sentinel_index]) * (max_other_sparse_effect + 0.01)
    }
  }

  # Combined sparse effects
  sparse_indices <- c(sentinel_index, other_sparse_indices)
  sparse_effects <- X[, sparse_indices] %*% beta[sparse_indices]

  # Scale sparse effects to achieve desired heritability
  scaling_factor_sparse <- sqrt(h2_sparse / var(sparse_effects))
  beta[sparse_indices] <- beta[sparse_indices] * as.vector(scaling_factor_sparse)

  # Ensure the sentinel SNP has the largest effect
  max_other_sparse_effect <- max(abs(beta[other_sparse_indices]))
  if (abs(beta[sentinel_index]) <= max_other_sparse_effect) {
    beta[sentinel_index] <- sign(beta[sentinel_index]) * (max_other_sparse_effect + 0.01)
  }

  # Oligogenic effects (adjust mixture proportions and sds)
  non_sparse_indices <- setdiff(1:n_features, c(sentinel_index, other_sparse_indices))
  n_oligogenic <- min(n_oligogenic, length(non_sparse_indices))
  oligogenic_indices <- sample(non_sparse_indices, n_oligogenic, replace = FALSE)

  mixture_assignments <- sample(1:length(mixture_props), length(oligogenic_indices), replace = TRUE, prob = mixture_props)
  beta[oligogenic_indices] <- rnorm(length(oligogenic_indices), 0, mixture_sds[mixture_assignments])

  # Scale oligogenic effects to achieve desired heritability
  oligogenic_effects <- X[, oligogenic_indices] %*% beta[oligogenic_indices]
  scaling_factor <- sqrt(h2_oligogenic / var(oligogenic_effects))
  beta[oligogenic_indices] <- beta[oligogenic_indices] * as.vector(scaling_factor)

  # Infinitesimal effects (small effects on remaining SNPs)
  infinitesimal_indices <- setdiff(non_sparse_indices, oligogenic_indices)
  infinitesimal_effects <- rep(0, n_features)
  if (length(infinitesimal_indices) > 0) {
    infinitesimal_effects[infinitesimal_indices] <- rnorm(length(infinitesimal_indices), 0, sqrt(h2_infinitesimal / length(infinitesimal_indices)))
  }
  beta <- beta + as.vector(infinitesimal_effects)

  # Generate y
  y <- X %*% beta

  # Add noise to achieve desired total heritability
  var_y <- var(as.vector(y))
  var_epsilon <- var_y * (1 - h2_total) / h2_total
  epsilon <- rnorm(n_samples, 0, sqrt(var_epsilon))
  y <- y + epsilon

  # Calculate actual heritabilities
  var_y_total <- var(as.vector(y))
  h2_sentinel_actual <- var(X[, sentinel_index] * beta[sentinel_index]) / var_y_total

  # Sparse effects heritability
  sparse_indices <- c(sentinel_index, other_sparse_indices)
  h2_sparse_actual <- var(X[, sparse_indices] %*% beta[sparse_indices]) / var_y_total

  # Oligogenic effects heritability
  h2_oligogenic_actual <- var(X[, oligogenic_indices] %*% beta[oligogenic_indices]) / var_y_total

  # Infinitesimal effects heritability
  h2_infinitesimal_actual <- var(X %*% infinitesimal_effects) / var_y_total

  h2_total_actual <- var(as.vector(X %*% beta)) / var_y_total

  ori.y <- y
  y <- scale(y, center = TRUE, scale = FALSE)

  # Create a full-length mixture_assignments vector
  mixture_assignments_full <- rep(NA, n_features)
  mixture_assignments_full[oligogenic_indices] <- mixture_assignments

  return(list(
    ori.X = ori.X,
    X = X,
    ori.y = ori.y,
    y = y,
    beta = beta,
    h2_total = h2_total_actual,
    h2_sparse = h2_sparse_actual,
    h2_sentinel = h2_sentinel_actual,
    h2_oligogenic = h2_oligogenic_actual,
    h2_infinitesimal = h2_infinitesimal_actual,
    sentinel_index = sentinel_index,
    other_sparse_indices = other_sparse_indices,
    oligogenic_indices = oligogenic_indices,
    infinitesimal_indices = infinitesimal_indices,
    mixture_assignments = mixture_assignments_full,
    var_epsilon = var_epsilon,
    sparse_indices = sparse_indices
  ))
}

# Function to Identify Causal SNPs Based on PVE Threshold
is_causal <- function(eqtl_data, pve_threshold) {
  # Get the beta vector and residual variance for this simulation
  beta <- eqtl_data$beta
  var_epsilon <- eqtl_data$var_epsilon

  # Compute variance explained by each SNP (since Var(X_j) = 1)
  variance_explained <- beta^2

  # Compute total genetic variance
  var_g <- sum(variance_explained)

  # Compute total variance (genetic variance + residual variance)
  total_variance <- var_g + var_epsilon

  # Compute PVE for each SNP
  proportion_var_explained <- variance_explained / total_variance

  # Define causal SNPs based on the current PVE threshold
  causal_SNPs <- which(proportion_var_explained > pve_threshold)
  return(causal = causal_SNPs)
}

generate_sparse_eqtl_data <- function(X,
                                      K = 10,      # Number of effect SNPs
                                      h2 = 0.3,    # Variance explained
                                      seed = NULL) {
  if (!is.null(seed)) set.seed(seed)

  n_samples <- nrow(X)
  n_features <- ncol(X)

  # Initialize beta (effect sizes) to zero
  beta <- rep(0, n_features)

  # (a) Sample causal SNP indices
  causal_indices <- sample(1:n_features, K, replace = FALSE)

  # (b) Assign N(0, 0.6^2) effects to causal SNPs
  beta[causal_indices] <- rnorm(K, mean = 0, sd = 0.6)

  # Compute genetic effects
  genetic_effect <- X %*% beta

  # (c) Solve for σ²
  var_g <- var(genetic_effect)  # Variance from genetic effects
  var_e <- var_g * (1 - h2) / h2  # Residual variance

  # (d) Generate phenotype y ~ N(Xβ, σ²)
  y <- genetic_effect + rnorm(n_samples, mean = 0, sd = sqrt(var_e))

  # Return results
  return(list(
    X = X,
    y = y,
    beta = beta,
    causal_indices = causal_indices,
    var_epsilon = var_e,
    h2_input = h2,
    h2_estimated = var_g / (var_g + var_e)
  ))
}
```


This vignette demonstrates an issue where our SuSiE.ash (Marginal) model fails to update its coefficients after the first iteration of SuSiE. In this simulation, we generate an oligogenic data setting and run the following steps:

1. Run a single iteration of SuSiE.

2. Extract the residuals.

3. Perform univariate regression on the residuals and compute & plot the z-scores to demonstrate there are signals SuSiE missed in the first iteration.

4. Continue on with the SuSiE.ash (Marginal) algorithm, specifically showcasing how the mr.ash component misses all signals. 

## Univariate Regression on Residuals

```{r, message = F, warning = F}
# Generate Data
data <- generate_eqtl_data(X = X20, h2_total = 0.3, mixture_props = c(0.75, 0.25), seed = 1)
data$causal <- is_causal(data, 0.005)

# Run single iteration of SuSiE
res_susie <- susie(X = data$ori.X, y = data$ori.y, L = 10, max_iter = 1, standardize = T, intercept = T)

# Extract Intercept + Coefficients
coefs         <- coef(res_susie)
intercept_hat <- coefs[1]
beta_hat      <- coefs[-1]

# Calculate Residuals
resid <- data$ori.y - (intercept_hat + data$ori.X %*% beta_hat)

# Run univariate regression on residuals + calculate z-scores
res_reg <- univariate_regression(X = data$ori.X, y = resid, center = F, scale = F)
z_scores <- res_reg$betahat / res_reg$sebetahat

# Calculate p-value + bonferroni adjustment
pvals <- 2 * pnorm(abs(z_scores), lower.tail = FALSE)
pvals_bonf <- p.adjust(pvals, method = "bonferroni")
sig_idx <- which(pvals_bonf < 0.05)  # no variants meet threshold
alpha_bonf <- 0.05 / length(pvals_bonf)
z_thresh   <- qnorm(1 - alpha_bonf / 2)
highlight_idx <- c(data$oligogenic_indices)

# Plot z-scores, highlight oligogenic indices, and include bonferonni threshold
plot(z_scores, main = "Univariate z-scores on the residuals\nPurple points denote oligogenic variants", ylab = "z-score", ylim = c(-5,5))
abline(h =  z_thresh, col = "red", lty = 2)
abline(h = -z_thresh, col = "red", lty = 2)
points(highlight_idx, z_scores[highlight_idx], col = "purple", pch = 19)
```

After the first iteration of SuSiE, none of the z-scores surpasses the Bonferroni threshold, although many variants (particularly oligogenic variants) come close. This suggests that there should still be a sufficient signals remaining in after the first iteration of SuSiE for the mr.ash component to capture. 

## Running the mr.ash update

```{r ash, message = F, include = F}
# Run first iteration of susie-ash marginal
res_marg <- susie_ash_RE_Marg(
                  X                = scale(data$ori.X),
                  y                = scale(data$ori.y),
                  L                = 10,
                  verbose          = F,
                  coverage         = 0.95,
                  XtX              = precomputed_matrices$XtX,
                  LD               = precomputed_matrices$LD,
                  V                = precomputed_matrices$V,
                  Dsq              = precomputed_matrices$Dsq,
                  VtXt             = precomputed_matrices$VtXt,
                  update_ash_sigma = F,
                  K.length         = 20,
                  upper_bound      = 2,
                  maxiter          = 1)
```


```{r plot}
# Compute the desired values
max_theta <- max(abs(res_marg$theta))
var_theta <- var(res_marg$theta)

plot(res_marg$theta, ylim = c(-1e-5, 1e-5),
     main = "Distribution of Theta Values after First\nIteration of SuSiE-ash (Marginal)",
     ylab = "Theta")
legend("topright", 
       legend = c(
         bquote(Max("|" * bold(theta) * "|") == .(format(max_theta, digits = 3))),
         bquote(Var(bold(theta)) == .(format(var_theta, digits = 3)))
       ),
       bty = "n")
```

The mr.ash coefficient values, theta, are very small--essentially negligible--despite there being sufficient signals after the first SuSiE iteration.   
